### 2004-07-16 Documentation/Manual.tex
part 1 finis

% Set as GLoBES command and add to index:
% Always use GLB when mentioning a command in text!
\newcommand{\GLB}[1]{{\tt glb#1}\index{api}{#1@{\tt #1}}}
% Always use GLBNS when adding a command to the index without showing it!
\newcommand{\GLBNS}[1]{\index{api}{#1@{\tt \protect{#1}}}}

% Set as GLoBES command and add to index:
% Always use GLB when mentioning a command in text!
\newcommand{\GLBC}[1]{{\tt #1}\index{constants}{\protect{\tt{#1}}}}
% Always use GLBNS when adding a command to the index without showing it!
\newcommand{\GLBNSC}[1]{\index{constants}{{\protect{\tt #1}}}}
\newcommand{\gq}[1]{\begin{quote} {\tt #1} \end{quote}}
\newcommand{\go}[1]{$\rightarrow$ Output: \begin{quote} {\sf #1} \end{quote}}
\newcommand{\bugs}{{\tt globes@ph.tum.de}}
\definecolor{light}{gray}{0.93}
\definecolor{heavy}{gray}{0.35}
\setlength{\fboxrule}{0.1cm}
\newcommand{\example}[2]{
\begin{table}[p]
\begin{center}
\fcolorbox{heavy}{light}{
\begin{minipage}[t][\textheight][t]{15.5cm}
{\em \underline{Example:} #1}

@@

\pagenumbering{Roman}
\thispagestyle{empty}
{
\setlength{\parindent}{0cm}

@@

\clearpage
\thispagestyle{empty}
\bigskip
\begin{quote}
    Copyright \copyright  2004  The GLoBES Team.
    Permission is granted to copy, distribute and/or modify this document
    under the terms of the GNU Free Documentation License, Version 1.2
    or any later version published by the Free Software Foundation;
    with no Invariant Sections, no Front-Cover Texts, and no Back-Cover Texts.
    A copy of the license is included in the section entitled "GNU
    Free Documentation License".
\end{quote}
\bigskip

@@

\GLOBES\ (``General Long Baseline Experiment Simulator'') is a flexible
software package to simulate neutrino oscillation 
long baseline experiments. On the
one hand, it contains a comprehensive abstract experiment definition
language, which allows to describe most classes of long baseline experiments
at an abstract level. On the other hand, it provides a C-library to 
process the experiment information in order to obtain oscillation
probabilities, rate vectors, and $\Delta \chi^2$-values. Currently, 
\GLOBES\ is available for Linux. Since the source code is included,
the modifications to other operating systems should be doable.

@@

@@

In \Part~\ref{part:2} of the manual, \AEDL\ is described. After an
introductory chapter, all functions are defined in greater detail.
This part might be more interesting for the experimental users who
want to modify or create \AEDL\ files. A useful tool in this context
is the software program ``globes'', which returns event rates and other
information for individual \AEDL\ files without further programming. 
For example, flux normalizations can with this tool be easily adjusted 
to reproduce the event rates of a specific experiment. It is described
in the last chapter of \Part~\ref{part:2}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% PART I: User's manual
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

@@

\printindex{api}{API functions}
\printindex{constants}{API constants \& macros}
\printindex{aedl}{AEDL reference}
\printindex{norm}{Index}

>>>>

% Set as GLoBES command and add to index:
% Always use GLB when mentioning a command in text!
\newcommand{\GLB}[1]{{\tt glb#1}\index{api}{#1@{\tt #1}}}
% Always use GLBNS when adding a command to the index without showing it!
\newcommand{\GLBNS}[1]{\index{api}{#1@{\tt #1}}}

% Set as GLoBES command and add to index:
% Always use GLB when mentioning a command in text!
\newcommand{\GLBC}[1]{{\tt #1}\index{constants}{\tt{#1}}}
% Always use GLBNS when adding a command to the index without showing it!
\newcommand{\GLBNSC}[1]{\index{constants}{{\tt #1}}}
\newcommand{\gq}[1]{\begin{quote} {\tt #1} \end{quote}}
\newcommand{\go}[1]{$\rightarrow$ Output: \begin{quote} {\sf #1} \end{quote}}
\newcommand{\bugs}{{\tt globes@ph.tum.de}}
\definecolor{light}{gray}{0.93}
\definecolor{heavy}{gray}{0.35}
\setlength{\fboxrule}{0.1cm}
\newcommand{\example}[2]{
\begin{table}[p]
\begin{center}
\fcolorbox{heavy}{light}{
\begin{minipage}[t][\textheight][t]{15.5cm}
{\em \underline{Example:} #1}

@@

\frontmatter
\pagenumbering{Roman}
\thispagestyle{empty}
{
\setlength{\parindent}{0cm}

@@

\clearpage
\thispagestyle{empty}
\bigskip
\begin{quote}
    Copyright \copyright  2004  The GLoBES Team.
    Permission is granted to copy, distribute and/or modify this document
    under the terms of the GNU Free Documentation License, Version 1.2
    or any later version published by the Free Software Foundation;
    with the invariant Sections ``Terms of usage of \GLOBES" 
    and ``Acknowledgments'', no Front-Cover Texts, and no Back-Cover Texts.
    A copy of the license is included in the section entitled "GNU
    Free Documentation License".
\end{quote}
\bigskip

@@

\GLOBES\ (``General Long Baseline Experiment Simulator'') is a flexible
software package to simulate neutrino oscillation 
long baseline and reactor experiments. On the
one hand, it contains a comprehensive abstract experiment definition
language (AEDL), which allows to describe most classes 
of long baseline experiments
at an abstract level. On the other hand, it provides a C-library to 
process the experiment information in order to obtain oscillation
probabilities, rate vectors, and $\Delta \chi^2$-values. Currently, 
\GLOBES\ is available for GNU/Linux. Since the source code is included,
the port to other operating systems is in principle possible.

@@

\chapter*{Terms of usage of \GLOBES}

\GLOBES\ is mainly developed for academic use. Thus The \GLOBES\ Team would
appreciate being given academic credit for it. Whenever you use \GLOBES\
to produce a publication or a talk indicate that you have used \GLOBES\ and
please cite the following reference
\begin{quote}
P. Huber, M. Lindner and W. Winter ... 
\end{quote}
and \emph{not} this manual. This manual does not constitute a scientific 
publication and will not be submitted to any scientific journal. 
It will evolve during time since it is intendend for 
regular revision.
Besides, many of the data which is used by \GLOBES\ and distributed together
with it should be correctly referenced where possible. For details see the
???.

Apart from that, \GLOBES\ is free software and open source, as such it is 
licensed under the GNU Public License.

@@

In \Part~\ref{part:2} of the manual, \AEDL\ is described. After an
introductory chapter, all functions are defined in greater detail.
This part might be more interesting for the experimental users who
want to modify or create \AEDL\ files. A useful tool in this context
is the software program \verb+globes+, which returns event rates and other
information for individual \AEDL\ files without further programming. 
For example, flux normalizations can with this tool be easily adjusted 
to reproduce the event rates of a specific experiment. It is described
in the last chapter of \Part~\ref{part:2}.

\paragraph{Note:} All examples for application software in C do require
a C++ compiler to be properly compiled. For pedagogical reasons, variable
declarations are done at that place where the variable is needed for the
first time, which is at variance with C syntax but not with C++ syntax.
That is the only way in which the examples deviate form ISO C.

@@

\mainmatter
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% PART I: User's manual
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\backmatter
\chapter*{Acknowledgments}
\addcontentsline{toc}{chapter}{Acknowledgments}

We would like to thank Martin Freund, who wrote the very first
version of a three-flavour matter profile treatment many years ago.
PH is especially thankful for the invaluable advice of Thomas Fischbacher on
many design issues. 
In addition, we would like to thank
Mark Rolinec for his help to translate the experiment descriptions
into \AEDL , and for creating the illustrations in this manuscript.
Finally, thanks to all the people who have been pushing this project
for many years, to the ones who have been continuing asking for the 
publication of the software, and the referees of several of our
papers for suggestions which lead to improvements in the software.

@@

\chapter{Indices}
\printindex{api}{API functions}

\printindex{constants}{API constants \& macros}

\printindex{aedl}{AEDL reference}

\printindex{norm}{Index}

### 2004-07-16 Documentation/Paper/paper.tex
added the one missing reference
part 1 finis

Finally, one ore more rules add up to an experiment, where the
total $\Delta \chi^2$ is obtained as the sum of the $\Delta \chi^2$
of all rules. This approach allows the definition of 
appearance and disappearance channels, neutrino and antineutrino running, 
or interaction types with different systematics (spectral information
versus counting rate) within one experiment. The \GLOBES\ user interface
allows the simulation of one or more experiment simultaneously, which
means that one could also use different experiments for different
oscillation channels. However, there is still one component missing
in the experiment definition, which is the matter density profile.
For an experiment, an arbitrary matter density profiles can be specified, 
which is evaluated with the evolution operator method REFERENCES.
In addition, the matter density profile is multiplied by a scaling factor, 
which is treated as an independent parameter
with a relative (matter density) uncertainty. With this
approach, one can take into account that the matter density profile along a 
specific baseline is only known to about $3\%-5\%$.

@@

For the future it is planed to further extend the flexibility of
\GLOBES\ by providing support for dynamically loadable modules, to
allow completely user defined systematics or non-standard oscillation
probabilities.

>>>>

Finally, one ore more rules add up to an experiment, where the
total $\Delta \chi^2$ is obtained as the sum of the $\Delta \chi^2$
of all rules. This approach allows the definition of 
appearance and disappearance channels, neutrino and antineutrino running, 
or interaction types with different systematics (spectral information
versus counting rate) within one experiment. The \GLOBES\ user interface
allows the simulation of one or more experiment simultaneously, which
means that one could also use different experiments for different
oscillation channels. However, there is still one component missing
in the experiment definition, which is the matter density profile.
For an experiment, an arbitrary matter density profiles can be specified, 
which is evaluated with the evolution operator method~\cite{Ohlsson:1999um}.
In addition, the matter density profile is multiplied by a scaling factor, 
which is treated as an independent parameter
with a relative (matter density) uncertainty. With this
approach, one can take into account that the matter density profile along a 
specific baseline is only known to about $3\%-5\%$.

@@

For the future it is planed to further extend the flexibility of
\GLOBES\ by providing support for dynamically loadable modules, to
allow completely user defined systematics or non-standard oscillation
probabilities. Additionally we plan to provide interfaces to various script
languages like Perl.
